### 2015-03-13 manuscript/plos_template.tex
work on manuscript

\subsection*{Application to pan-cancer somatic mutation data}

Finally, we have applied the proposed method for the pan-cancer somatic mutation data \cite{pmid23945592}.
For each cancer type, the proposed method was applied separately.
By changing the number of mutation signatures and calculating values of the log-likelihood and bootstrap errors,
we have chosen the number of mutation signatures for each cancer type.
The estimated mutation signatures across cancer types were clustered by Frobenius Distance.
 % (see Supplementary Material for the values of the log-likelihoods and bootstrap errors for each cancer type).

@@

The signature 13 (T $>$ [AGT] at TpCpN sites) was observed in 12 cancer types, and related to the activity of APOBEC family.
The composition of nucleotides in estimated signatures in 12 different cancer types are highly consistent (Figure \ref{APOBEC_sig_list}) except for the on in B-cell lymphoma,
and almost of of them showed slightly higher abundance of A and T bases and depletion of G base at $-2$ positions (Figure \ref{APOBEC_twofive}).
This fact backs up the results obtained from UCUT data in the previous subsection and the previous study \cite{pmid23318258},
and implies the robustness of the proposed approach.
In terms of transcribed strand-specificity, the tendency is not consistent, implying that there is no significant strand-specificity in APOBEC signatures.

The signature 1 and 8 (C $>$ A at TpCpT and C $>$ T at TpCpG, respectively) were observed in colorectal and uterine cancers,
are probably associated with deregulated activity of the error-prone polymerase Pol $\epsilon$.
The proposed method used two signatures for representing the signature for the Pol $\epsilon$ dysfunction, 
while was represented by one signature in the previous study.
Interestingly, signature 1 showed transcription biases whereas signature 8 did not.
This phenomena are consistent for both colorectal and uterus cancers (Figures \ref{POLE1_sig_list} and \ref{POLE2_sig_list}).
C $>$A at TpCpT mutations were shown to be enriched in leading strands of replication forks.
whereas, there is no mention about the replication strand-specificity of C $>$ T at TpCpG mutations \cite{pmid25228659}.
Although replication strand biases is different from the transcription strand biases, these two biases may have strong corelation,
considering the fact that replication origins prefer transcription start sites \cite{pmid23187890}.
Therefore, the transcription strand bias observed in the signature 1 may actually come from replication strand biases.

Furthermore, both the signature 1 and 8 signatures showed strong abundance of T base at $-2$ positions,
and abundance of T bases at two bases 5' and 3' to the mutated site, respectively (Figures \ref{POLE1_sig_list} and \ref{POLE2_sig_list}).
In fact, nonsense mutations in TP53 at R231X, which are shown to be enriched in cancers having Pol $\epsilon$,
and abundance of T bases at two bases 5' and 3' to the mutated site, respectively (Figures \ref{POLE1_sig_list} and \ref{POLE2_sig_list}).
In the previous study of Pol $\epsilon$, a nonsense mutation in TP53 at R23X were shown to be enriched in cancers with Pol $\epsilon$ defects.
In fact, this mutation is C $>$ T at TpTpCpGpA, the most likely patten with the signature 8.
This implies that observing $\pm 2$ bases may be helpful for identifying the cause of mutations with higher resolution.

The signature 18 (T $>$ C at ApTp[AG]), observed in liver cancer, shows very strong strains specificity,
were shown to be more common in Asian cases than in other ancestries \cite{pmid25362482}, though the source of this signature is still not clear.

The signature 3 and 5 (C $>$ A at NpCpN), which were observed in head-and-neck cancers, three types of lung cancers,
are probably associated with tobacco smoking.
Furthermore, the fact that the estimated signature for each cancer shows higher prevalence on coding strands is consistent with the previous study.

The signature 7 (C $>$ T at NpCpG sites) was observed in 25 out of 30 cancer type, and related to deamination of 5-methyl-cytosine.
The signature 9 (C $>$ T at NpCp[CT]) were observed in melanomas and glioblastomas, 
and seems to be associated with a chemotherapy drug, temozolomide.

The signature 9 (C $>$T at [CT]pCpC) were observed in head and neck cancers and melanomas, and probably related to ultraviolet light.
The signature 10 (C $>$ T at GpCp[CG]) were observed in small-cell lung cancers and stomach cancers,
and seems to be the same with the ``signature 15'' in the previous study, whose function is still not clear.

The signature 2 (C $>$ A at [CT]pCpT), observed in low grade gliomas, can be related to the one detected in the same cancer type in the previous study, but slightly different. 
This seems to be because the corresponding signature in the previous study shows very complex pattern (C $>$ A at NpCpT or C $>$ T at GpTpN).
Actually, the mutation related to this signature is mostly coming from one cancer with a very high mutation rate.
When we re-analyze the data removing this sample, the signature 2 disappeared.
We suspect that the low-grade-glioma specific signature detected in the previous study may be aggregation of multiple signatures,
but cannot be decomposed due to the lack of samples having these signatures.
Therefore, we need to be more careful for the signature 2 and the corresponding signature in the previous study.

The signature 3 (C $>$ A at CpCpG), observed in several cancers,
Actually, this signature correspond to the one that could not be validated in the previous study,
and the pattern

@@

Although, our method detected most of the

Many of the signatures captured in this study 
had prominent features at the two bases 5' and 3' to the mutated sites.
For the signature 13, the signatures captured in 11 different cancer types are highly consistent (Figure \ref{APOBEC_sig_list}),
and they consistently showed depletion of G base at two bases 5' to the mutated site (Figure \ref{APOBEC_twofive}),
backing up the result obtained from UCUT data in the previous subsection and the previous study  \cite{pmid23318258}.
For the Pol $\epsilon$ signatures (signature 7 and signature 1), 
signatures for the two different cancer types consistently showed abundance of T base at two bases 5' to the mutated site,
and abundance of T bases at two bases 5' and 3' to the mutated site, respectively (Figures \ref{POLE1_sig_list} and \ref{POLE2_sig_list}).
For the signature 19, depletion of C base at two bases 5' to the mutated site is consistently observed (Figure \ref{LMST_sig_list}).
Although, as far as we know, there was no literature showing the characteristics of mutation signatures at $\pm 2$ sites,
Considering from the robustness shown in the APOBEC signature, many of  the characteristics inferred here might be true.

@@

% \caption{sample}
% \label{label 2}
\end{flushleft}
\end{minipage}
\caption{An example of a mutation signature and its visualization in the proposed approach.
Here, mutation features (substitution patterns, two 5' and 3' bases and strand direction) 
are assumed to be independent ($L=6$, $\bm{M} = (6, 4, 4, 4, 4, 2)$).
In the bottom, the size of each box represents the frequencies of bases (A, C, G and T) at the flanking sites.
In the top, the height of each box represents the conditional frequencies of mutated bases for each original base (C and T).
In the upper right, the height of the $+$ (non-coding strand) and $-$ (coding strand) boxes represents the frequencies of  stranded direction of mutations.
}
\label{mutSig_example}

>>>>

\subsection*{Application to somatic mutation data of 30 cancer types}

Finally, we have applied the proposed method for the somatic mutation data of 30 cancer types \cite{pmid23945592}.
For each cancer type, the proposed method was applied separately to see the robustness of the estimated signatures among different cancer types.
By changing the number of mutation signatures and calculating values of the log-likelihood and bootstrap errors,
we have chosen the number of mutation signatures for each cancer type.
The estimated mutation signatures across cancer types were clustered by Frobenius Distance.
 % (see Supplementary Material for the values of the log-likelihoods and bootstrap errors for each cancer type).

@@

The signature 13 (T $>$ [AGT] at TpCpN sites) was observed in 12 cancer types, and surely related to the activity of APOBEC family.
The composition of nucleotides in estimated signatures in 12 different cancer types are highly consistent (Figure \ref{APOBEC_sig_list}) except for the on in B-cell lymphoma,
and almost of of them showed slightly higher abundance of A and T bases and depletion of G base at $-2$ positions (Figure \ref{APOBEC_twofive}).
This fact backs up the results obtained from UCUT data in the previous subsection and the previous study \cite{pmid23318258},
and implies the robustness of the proposed approach.
The estimated transcribed strand-specificities varied across cancer types, implying that there is no significant strand-specificity in APOBEC signatures.

The signature 1 and 8 (C $>$ A at TpCpT and C $>$ T at TpCpG, respectively) were observed in colorectal and uterine cancers,
are associated with deregulated activity of the error-prone polymerase Pol $\epsilon$.
The proposed method used two signatures for representing the signature for the Pol $\epsilon$ dysfunction, 
while was represented by one signature in the previous study.
Interestingly, signature 1 showed transcription biases whereas signature 8 did not.
This phenomena are consistent for both colorectal and uterus cancers (Figures \ref{POLE1_sig_list} and \ref{POLE2_sig_list}).
C $>$A at TpCpT mutations were shown to be enriched in leading strands of replication forks.
whereas, there is no mention about the replication strand-specificity of C $>$ T at TpCpG mutations \cite{pmid25228659}.
Although replication strand biases is different from the transcription strand biases, these two biases may have strong correlation,
considering the fact that replication origins prefer transcription start sites \cite{pmid23187890}.
Therefore, the transcription strand bias observed in the signature 1 may actually come from replication strand biases.

Furthermore, both the signature 1 and 8 signatures shows strong abundance of T base at $-2$ positions,
and abundance of T bases at two bases 5' and 3' to the mutated site, respectively (Figures \ref{POLE1_sig_list} and \ref{POLE2_sig_list}).
In fact, nonsense mutations in TP53 at R231X, which are shown to be enriched in cancers having Pol $\epsilon$,
and abundance of T bases at two bases 5' and 3' to the mutated site, respectively (Figures \ref{POLE1_sig_list} and \ref{POLE2_sig_list}).
In the previous study of Pol $\epsilon$, a nonsense mutation in TP53 at R23X were shown to be enriched in cancers with Pol $\epsilon$ defects.
In fact, this mutation is C $>$ T at TpTpCpGpA, the most likely patten with the signature 8.
This implies that observing $\pm 2$ bases may be helpful for identifying the cause of mutations with higher resolution.

The signature 2 (C $>$ A at [CT]pCpT), observed in low grade gliomas, can be related to the one detected in the same cancer type in the previous study, but slightly different. 
This seems to be because the corresponding signature in the previous study shows very complex pattern (C $>$ A at NpCpT or C $>$ T at GpTpN).
Actually, the mutation related to this signature is mostly coming from one sample with a very high mutation rate, 
and the signature 2 disappeared when we re-analyze the data removing this sample.
We suspect that the complex low-grade-glioma specific signature detected in the previous study may be aggregation of multiple signatures,
but cannot be decomposed due to the lack of the number of samples having these signatures.
Therefore, we need to be more careful for the signature 2 and the corresponding signature in the previous study.

The signature 3 and 5 (C $>$ A at NpCpN) observed in head-and-neck cancers and three types of lung cancers are probably associated with tobacco smoking.
The estimated signature in each cancer type shows higher mutation prevalence on the template strand, which is consistent with the previous study \cite{pmid12379884, pmid23318258}.

The signature 4 (C $>$ A at CpCpG) observed in several cancers shows remarkably high ratio of base C at the $-2$ position.
This seems to correspond to the ``signatures R2''  (detected in kidney clear cell carcinomas, lung adenocarcinomas, lung squamous carcinomas and melanomas) 
that could not be validated in the previous study (see supplementary figures of the previous study \cite{pmid23318258}),
meaning that significant ratios of somatic mutations corresponding to that signature was not estimated to be false positives by re-sequencing or visual inspection of BAM files using genomic viewers.
Actually, the pattern of this signature matches to that of artifact by oxidation of DNA during acoustic shearing \cite{pmid23303777}.
Although this signature may be from the artifacts, knowing higher resolution of the artifact signature is very helpful to accurately remove the false positive mutations.

The signature 18 (T $>$ C at ApTp[AG]) observed in liver cancer shows very strong strains specificity, implying the evidence of transcription-coupled nucleotide excision repair.
This signature has been shown to be more common in Asian cases than in other ancestries \cite{pmid25362482}, though the source of this signature is still not clear.

The signature 7 (C $>$ T at NpCpG sites) was observed in 25 out of 30 cancer type, and related to deamination of 5-methyl-cytosine.
The signature 9 (C $>$ T at NpCp[CT]) were observed in melanomas and glioblastomas, 
and are probably associated with a chemotherapy drug, temozolomide.
The signature 10 (C $>$T at [CT]pCpC) were observed in head and neck cancers and melanomas, and probably related to ultraviolet light.

The signature 11 (C $>$ T at GpCp[CG]) were observed in small-cell lung cancers and stomach cancers,
and seems to be the same with the ``signature 15'' in the previous study, whose function is still not clear.
When we investigate $\pm 2$ positions, the depletion of C base at two bases 5' to the mutated site is consistently observed (Figure \ref{LMST_sig_list}).

@@

@@

% \caption{sample}
% \label{label 2}
\end{flushleft}
\end{minipage}
\caption{An example of a mutation signature and its visualization in the proposed approach.
Here, mutation features (substitution patterns, two 5' and 3' bases and strand direction) 
are assumed to be independent ($L=6$, $\bm{M} = (6, 4, 4, 4, 4, 2)$).
In the bottom, the size of each box represents the frequencies of bases (A, C, G and T) at the flanking sites.
In the top, the height of each box represents the conditional frequencies of mutated bases for each original base (C and T).
In the upper right, the height of the $+$ box represents the frequencies of mutations in the coding strand (or the plus strand, the sense strand and the untranscribed strand)
whose nucleotide sequences directly corresponds to mRNA, 
whereas the height of $-$ box represents those in the template strand (or the minus strand, the antisense strand, the transcribed strand and the noncoding strand)
whose sequences are copied during the synthesis of mRNA.
}
\label{mutSig_example}
